---
title: "creating graphs with ggplot2"
author: "David Miller"
date: "1/27/2020"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse.

we will install these packages
-tidyverse: is a suite of packages for wrangling and viz (ggplot)
-readxl: to read excel files (in many ways)
-here: to help with filepaths
```{r setup, message=FALSE}
#library(tidyverse) # install.packages("tidyverse")
library(ggplot2)
library(readr)
library(readxl) # install
library(here) # install

```

#Cal National Parks data
```{r}
ca_np <- read_csv(here("data/ca_np.csv"))
# also be written
ca_np <- read_csv("data/ca_np.csv")

```

```{r}
ci_np <- read_excel(here("data", "ci_np.xlsx"))
```

# Make a plot

```{r}
ggplot(data=ci_np, aes(x = year, y = visitors)) +
  geom_contour()

```
Make a base layer for my CI plots:
```{r}
ggbase <- ggplot(data=ci_np, aes(x = year, y = visitors))
ggbase + geom_point()

```
### Now lets start customizing!

-line / point colors:  color
-change line type:  linetype
-change point style:  pch
-change transparency: alpha

```{r}
ggbase +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

Activity: create a ggplot scatterplot with triangle symbols (hint = ?pch), of a non-default size type (size = #), and that are an updated color

```{r}
gg_base = ggbase
ggbase+
  geom_point(
    color = "dodgerblue",
    pch = 2,
    size = 4
  )

  )
```

### Update graph aesthetics based on variables

```{r}
ca_np <- read.csv(here("data","ca_np.csv"))
```

ggplot 

ggbase + 
geom_point(
aes(size = visitors,
color = year)
)
```{r}
ggbase + 
geom_point(
aes(size = visitors,
color = year)
)

```
```{r}
gg_base +
  geom_area() +
  theme_gray()
```

# Add an arguement

##### Use gg-themes
- theme_light()
- theme_bw()


### Create a multi-series line plot using data for all CA National Parks

```{r}
ggplot(data = ca_np, aes(x=year, y=visitors, group =
                           park_name)) + 
  geom_line()
```

#Try to update the line colors based upon the park_name variable

```{r}
ggplot(data = ca_np, aes(x=year, y=visitors, group =
                           park_name)) + 
  geom_line(
    aes(color =park_name)
  )
```

```{r}
my_np_graph <- ggplot(data = ca_np, aes(x=year, y=visitors)) + 
  geom_point(
    aes(color =park_name), show.legend=FALSE) +
  facet_wrap(~park_name, scales = "free") +
  theme_bw()
my_np_graph
```

Export my graph as a .png
```{r}
ggsave(plot = my_np_graph, here("Figures", "np_graph.png"))
```

