---
title: "Tidying Session"
author: "David Miller"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# Attaching packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)
```


### Read in inverts.xlxs

```{r}
inverts <- 
  read_excel(here("data", "inverts.xlsx"))
inverts
```

###`pivot_longer()`
Use a single colon (:) to indicate a series of sequential columns A:D
```{r}
inverts_long <- pivot_longer(data=inverts,
                             cols = '2016':'2018',
                             names_to = "year",
                             values_to = "sp_count")
inverts_long

# what if want some in series and some not in series?
# c('2016':'2018', '2020')
```

what if want to go back to wide format?
```{r}
inverts_wide <- inverts_long %>%
  pivot_wider(names_from=common_name,
              values_from = sp_count)
inverts_wide
```

### `janitor::clean_names()`

```{r}
inverts_wide <- inverts_wide %>%
  clean_names
inverts_wide
```

### `tidyr::unite()` to combine information from columns

Example:merge the site code and year into a single column called "site_year"
```{r}
inverts_unite <- inverts_long %>%
  unite(col = "site_year",
        c(site, year),
        sep = "_")
inverts_unite
```

Create a new object called "inverts_moyr" starting from inverts_long that unites the month and year colums into a single column named "mo_yr" using a "/" as the seperator
```{r}
inverts_moyr <- inverts_long %>%
  unite(col = "mo_yr",
        c(month, year),
        sep = "/")
inverts_moyr
```

### `tidr::seperate`
```{r} 
inverts_sep <- inverts_unite %>%
  separate(col = site_year, into = c("my_site", "my_year"))
```

#add 5 to all tree heights
```{r}
trees_new <- trees %>%
  mutate(height_update = Height + 5)
```
### `str_replace`

```{r}
ca_abbr <- inverts %>%
  mutate(common_name = 
           str_replace(common_name,
                       pattern = "california",
                       replacement = "CA"))
```

###Example:  in the common name column you want to replace "spiny" with "awesome"

store as "awesome_lobsters"

```{r}
awesome_lobsters <- inverts %>%
  mutate(common_name = 
           str_replace(common_name,
                       pattern = "spiny",
                       replacement = "awesome"))
```

